Molecular alignment at graphene surface determines tansport properties of graphene transistors

### 2020-04-03 manuscript.tex
Update manuscript.tex
Transconductance of clean graphene
Transconductance - 1st coverage

how did you get 0.073 eV?

The understanding of morphological properties of the OS layer and its growth mode is important for the interpretation of the charge transport data presented in the following.  In organic semiconductors charge transport proceeds by hopping in a manifold of localized molecular electronic states and is thus strongly dependent on the relative orientation of the molecules and the degree of positional disorder in the layers.  On graphene instead, charge transport is characterized by Bloch-like extended states, which reside, in the case of a single-layer graphene in a relatively disorder-free environment.  The effect of PDI8-CN2 layers on charge transport through graphene was studied by measuring the transfer characteristics of GFETs with different amounts of PDIF8-CN2 grown on graphene channel. Typical evolution of resistance for clean graphene and  two different coverages is illustrated in Figure \ref{fig:3}, where we show transfer the results of measurements (symbols) obtained on GFETs comprising clean single-layer (red curve), GFET with PDI8-CN2 coverage of 1 minute and PDI8-CN2 (blue curve) and with coverage of 3 minutes (green curve).  Only a forward gate sweep (from $V_g<0$ to $V_g>0$) data are shown. A minimal hysteresis was observed (See Supplementary information) indicating low density of traps that would originate from the structural and/or chemical defects in the layers. Also shown are the results of fitting to a model function (black solid curves) explained below.

\begin{figure}[htb]
  \centering
   \includegraphics[width=0.9\textwidth]{./Figures/fig3}  
  \caption{Transconductance measurements of a graphene field-effect transistor at different PDI8-CN2 coverages of graphene. Red curve corresponds to clean graphene, blue curve corresponds to a coverage of 0.6 obtained by the PDI8-CN2 deposition of 1 minute. The green curve corresponds to the coverage obtained by PDI8-CN2 deposition of 3 minute. The corresponding morphologies are shown in Figures 1 and 2, respectively. }
  \label{fig:3}
\end{figure}

We observe transfer characteristics of GFET comprising clean graphene exhibiting charge-neutrality point at back-gate voltage of 8.5 V.  The position of CNP relative to the Dirac point (0 V) is consistent with $p$-type doping of graphene caused by the trap states on SiO$_{2}$ surface and chemical redox reaction occurring at the graphene/SiO$_{2}$ interface\cite{feng-2014,peng-2017,lafkioti-2010}.  In addition to the shift of the conductivity minimum we observe also a strong reduction in the slope of the transfer curve occurring at about 3.6 V, leading to a plateau of the drain current.  Such behavior is commonly present in GFETs with relatively short channels, where the role of the metallic contacts becomes important\cite{bartolomeo-2015,nouchi-2008,nouchi-2014}. Two principal mechanisms in GFET affect charge transport under ambient conditions. First is already mentioned doping of the channel with airborne species. The second is related to the graphene doping under metallic contacts\cite{farmer-2009,giovannetti-2008,lee-2008,mueller-2009}.  Several groups report the evidence that metallic contacts induce doping of the underlying graphene \cite{giovannetti-2008,mueller-2009,huard-2008}, and the amount of transferred charge depends on the difference of metallic work function and the work function of graphene.  Specifically, Au contacts used in our experiments cause p-type doping, according to the work by Giovanetti et al.\cite{giovannetti-2008}, causing a shift in Fermi level of 0.3 eV in the direction away from the vacuum level, across the Dirac point. In addition, changes in the interface charge density may be a consequence of Fermi-level depinning at the metal/graphene interface and is related to microscopic interface environment that is characterized by interdiffusion of metallic atoms and/or oxide interlayer\cite{nouchi-2008,nouchi-2014}. Since Fermi level at such interfaces is not pinned or only partially pinned\cite{dibartolomeo-2011}, the interface charge density may be affected by the applied gate voltage, which is not the case for the abrupt interfaces.  In the channel instead, we can calculate the position of the Fermi level relative to the Dirac point in graphene ($\Delta E$) from the measured position of CNP (Figure \ref{fig:3}), using the approach from Ref. \onlinecite{wang-2008}:
\begin{equation}
  \label{eq:1}
  \Delta E = \hbar v_{F}\sqrt{\pi\alpha|V_{g}-V_{D}}
\end{equation}

where $v_F=0.8\times10^6$ m/s, and $\alpha=7.2\times10^{14}$ cm$^{-2}$ V$^{-1}$. For the value $|V_g-V_D |=8.5$ V we obtain the position of the Fermi level to be 0.073 eV below the Dirac point. The channel structure can therefore be regarded as $p^{++}/p/p^{++}$. A consequence of such variation of charge density within the structure are deviations from the linear transfer lineshapes, which in extreme cases may result in curves may extreme with double dips\cite{nouchi-2014,nam-2012}, or in less extreme exhibiting an abrupt change in slope continuing to a plateau.

@@

At the PDI8-CN2 coverage of 0.6 the position of CNP is shifted slightly (for ~0.5 V)  to more positive voltages and the current at the CNP is slightly increased relative to the clean graphene.  More importantly, we note a less pronounced change in the slope of the transfer characteristics at negative gate voltages, compared to the one observed on clean graphene (red curve). The observed behavior of the transport properties can be understood in the context of rather complex morphology of PDI8-CN2 layers on graphene, and with is associated interaction of the PDI8-CN2 molecules with graphene. At the sub monolayer coverage of PDI8-CN2 the hole transport continues to proceed through the channel, relatively unhindered, electron transport, however is enhanced relative to clean graphene (cf. red and blue curve in Figure \ref{fig:3}). We have noted that the crossover in the transfer curve is related to the partial depinning of the Fermi level near the contacts. Due to the low density of states in graphene this interface environment may extend as far as 0.2 $\mu$m\cite{mueller-2009}. This is the area exposed to the flux of incoming PDI8-CN2 molecules during the organic layer growth and represents a place of nucleation of the PDI8-CN2 islands. The AFM results (Figure 1) show that at this coverage islands are formed in which the molecules are oriented face-on on the graphene surface. It has been shown\cite{su-2009} that molecules with large aromatic structures carrying surplus of negative charge are noncovalently bonded to graphene surface by $\pi-\pi$ interactions, which result in strong charge transfer at the interface. The additional charges that are brought to the interface by the organic molecules change the carrier concertation in graphene and change the position of the Fermi level, in the direction as to reduce the $p^{++}$ doping at the interface and therefore increase hole mobility in that region. The low position of LUMO in PDI8-CN2 (-4.3 eV)\cite{jones-2007} enables efficient electron transfer to graphene, reducing strong p-type doping levels near the Au contacts. Consequently, the cusp in the transfer curve is reduced.   As for the small shift of CNP relative to the clean graphene, it corresponds to to the change of position of Fermi level of only ~$3\times10^{-3}$ eV, indicating negligible charge transfer between PDI8-CN2 layer and graphene, which would affect its electronic properties. Interestingly, the 0.6 monolayer coverage agrees nicely with the 0.5 ML--0.7ML coverage of PDI8-CN2 at which the onset of  drain-source current is observed in PDI8-CN2-based organic thin film transistors\cite{liscio-2013}.

@@

Qualitatively, we can apply these arguments also for the case of PDI8-CN2 molecules. In the absence of available calculations of the electrostatic potential over the PDI8-CN2 molecule, we can speculate that the imide groups provide an electric dipole that points away from the surface and therefore increases electron binding energy. On the other hand DFT calculations of the electrostatic potential of the perylene core by Geng et al.\cite{geng-2012} show an abundance of negative charge inside the core, implying  that the $\pi$-electron system above the molecule represents an electron-deficient region, resulting in a surface dipole pointing towards the surface. Consequently, the electron binding energy in the layer of flat-lying molecules is lower relative to the upright oriented molecules.  In the case of 4 minute exposure the energy band alignment in the resulting structure corresponds to a type-II heterostructure with graphene and the intermediate layer of flat-lying molecules with energy band structure shifted closer to the vaccum level than the outer layer.  The direction of the dipole electric field of the first molecular layer precludes appreciable induction of the electric charge due to the back-gate, and slightly shifts CNP point towards the positive gate voltages. As the second molecular layer is completed, the resulting dipole electric field -- directed away from the surface favors electron accumulation at the graphene interface resulting in a strong shift of the CNP in the negative direction.  The molecular density of the second layer is considerably higher with perylene core interplanar of $< 0.34$ nm, which results in higher dipole electric field intensity than in  case of the layer comprising the flat-lying molecules.

>>>>

The understanding of morphological properties of the OS layer and its growth mode is important for the interpretation of the charge transport data presented in the following.  In organic semiconductors charge transport proceeds by hopping in a manifold of localized molecular electronic states and is thus strongly dependent on the relative orientation of the molecules and the degree of positional disorder in the layers.  On graphene instead, charge transport is characterized by Bloch-like extended states, which reside, in the case of a single-layer graphene in a relatively disorder-free environment.  The effect of PDI8-CN2 layers on charge transport through graphene was studied by measuring the transfer characteristics of GFETs with different amounts of PDI8-CN2 grown on graphene channel. Typical evolution of resistance for clean graphene and  two different coverages is illustrated in Figure \ref{fig:3}, where we show transfer the results of measurements (symbols) obtained on GFETs comprising clean single-layer (brown curve), GFET with PDI8-CN2 coverage of 1 minute and PDI8-CN2 (blue curve) and with coverage of 3 minutes (green curve).  Only a forward gate sweep (from $V_g<0$ to $V_g>0$) data are shown. A minimal hysteresis was observed (See Supplementary information) indicating low density of traps that would originate from the structural and/or chemical defects in the layers. Also shown are the results of fitting to a model function (black solid curves) explained below.

\begin{figure}[htb]
  \centering
   \includegraphics[width=0.9\textwidth]{./Figures/fig3}  
  \caption{Transconductance measurements of a graphene field-effect transistor at different PDI8-CN2 coverages of graphene. Brown curve corresponds to clean graphene, blue curve corresponds to a coverage of 0.7 obtained by the PDI8-CN2 deposition of 1 minute. The green curve corresponds to the coverage obtained by PDI8-CN2 deposition of 4 minutes. The corresponding morphologies are shown in Figures 1 and 2, respectively. }
  \label{fig:3}
\end{figure}

We observe transfer characteristics of GFET comprising clean graphene exhibiting charge-neutrality point at back-gate voltage of 8 V.  The position of Charge Neutrality Point (CNP) relative to the Dirac point (0 V) is consistent with $p$-type doping of graphene caused by the trap states on SiO$_{2}$ surface and chemical redox reaction occurring at the graphene/SiO$_{2}$ interface\cite{feng-2014,peng-2017,lafkioti-2010}.  In addition to the shift of the conductivity minimum we observe also a strong reduction in the slope of the transfer curve occurring at about 1 V, leading to a plateau of the drain current.  Such behavior is commonly present in GFETs with relatively short channels, where the role of the metallic contacts becomes important\cite{bartolomeo-2015,nouchi-2008,nouchi-2014}. Two principal mechanisms in GFET affect charge transport under ambient conditions. First is already mentioned doping of the channel with airborne species. The second is related to the graphene doping under metallic contacts\cite{farmer-2009,giovannetti-2008,lee-2008,mueller-2009}.  Several groups report the evidence that metallic contacts induce doping of the underlying graphene \cite{giovannetti-2008,mueller-2009,huard-2008}, and the amount of transferred charge depends on the difference of metallic work function and the work function of graphene.  Specifically, Au contacts used in our experiments cause p-type doping, according to the work by Giovanetti et al.\cite{giovannetti-2008}, causing a shift in Fermi level of 0.3 eV in the direction away from the vacuum level, across the Dirac point. In addition, changes in the interface charge density may be a consequence of Fermi-level depinning at the metal/graphene interface and is related to microscopic interface environment that is characterized by interdiffusion of metallic atoms and/or oxide interlayer\cite{nouchi-2008,nouchi-2014}. Since Fermi level at such interfaces is not pinned or only partially pinned\cite{dibartolomeo-2011}, the interface charge density may be affected by the applied gate voltage, which is not the case for the abrupt interfaces.  In the channel instead, we can calculate the position of the Fermi level relative to the Dirac point in graphene ($\Delta E$) from the measured position of CNP (Figure \ref{fig:3}), using the approach from Ref. \onlinecite{wang-2008}:
\begin{equation}
  \label{eq:1}
  \Delta E = \hbar v_{F}\sqrt{\pi\alpha|V_{g}-V_{D}|}
\end{equation}

where $v_F=0.8\times10^6$ m/s, and $\alpha=7.2\times10^{14}$ cm$^{-2}$ V$^{-1}$. For the value $|V_g-V_D |=8$ V we obtain the position of the Fermi level to be 0.073 eV below the Dirac point. The channel structure can therefore be regarded as $p^{++}/p/p^{++}$. A consequence of such variation of charge density within the structure are deviations from the linear transfer lineshapes, which in extreme cases may result in curves may extreme with double dips\cite{nouchi-2014,nam-2012}, or in less extreme exhibiting an abrupt change in slope continuing to a plateau.

@@

At the PDI8-CN2 coverage of 0.7 the position of CNP is shifted slightly (for ~0.5 V)  to more positive voltages and the current at the CNP is slightly increased relative to the clean graphene.  More importantly, we note a less pronounced change in the slope of the transfer characteristics at negative gate voltages, compared to the one observed on clean graphene (brown curve). The observed behavior of the transport properties can be understood in the context of rather complex morphology of PDI8-CN2 layers on graphene, and with is associated interaction of the PDI8-CN2 molecules with graphene. At the sub-monolayer coverage of PDI8-CN2 the hole transport continues to proceed through the channel, relatively unhindered, electron transport, however is enhanced relative to clean graphene (cf. brown and blue curve in Figure \ref{fig:3}). We have noted that the crossover in the transfer curve is related to the partial depinning of the Fermi level near the contacts. Due to the low density of states in graphene this interface environment may extend as far as 0.2 $\mu$m\cite{mueller-2009}. This is the area exposed to the flux of incoming PDI8-CN2 molecules during the organic layer growth and represents a place of nucleation of the PDI8-CN2 islands. The AFM results (Figure 1) show that at this coverage islands are formed in which the molecules are oriented face-on on the graphene surface. It has been shown\cite{su-2009} that molecules with large aromatic structures carrying surplus of negative charge are noncovalently bonded to graphene surface by $\pi-\pi$ interactions, which result in strong charge transfer at the interface. The additional charges that are brought to the interface by the organic molecules change the carrier concertation in graphene and change the position of the Fermi level, in the direction as to reduce the $p^{++}$ doping at the interface and therefore increase hole mobility in that region. The low position of LUMO in PDI8-CN2 (-4.3 eV)\cite{jones-2007} enables efficient electron transfer to graphene, reducing strong p-type doping levels near the Au contacts. Consequently, the cusp in the transfer curve is reduced.   As for the small shift of CNP relative to the clean graphene, it corresponds to to the change of position of Fermi level of only ~$3\times10^{-3}$ eV, indicating negligible charge transfer between PDI8-CN2 layer and graphene, which would affect its electronic properties. Interestingly, the 0.7 monolayer coverage agrees nicely with the 0.5 ML--0.7ML coverage of PDI8-CN2 at which the onset of  drain-source current is observed in PDI8-CN2-based organic thin film transistors\cite{liscio-2013}.

@@

Qualitatively, we can apply these arguments also for the case of PDI8-CN2 molecules. In the absence of available calculations of the electrostatic potential over the PDI8-CN2 molecule, we can speculate that the imide groups provide an electric dipole that points away from the surface and therefore increases electron binding energy. On the other hand DFT calculations of the electrostatic potential of the perylene core by Geng et al.\cite{geng-2012} show an abundance of negative charge inside the core, implying  that the $\pi$-electron system above the molecule represents an electron-deficient region, resulting in a surface dipole pointing towards the surface. Consequently, the electron binding energy in the layer of flat-lying molecules is lower relative to the upright oriented molecules.  In the case of 4 minutes exposure the energy band alignment in the resulting structure corresponds to a type-II heterostructure with graphene and the intermediate layer of flat-lying molecules with energy band structure shifted closer to the vaccum level than the outer layer.  The direction of the dipole electric field of the first molecular layer precludes appreciable induction of the electric charge due to the back-gate, and slightly shifts CNP point towards the positive gate voltages. As the second molecular layer is completed, the resulting dipole electric field -- directed away from the surface favors electron accumulation at the graphene interface resulting in a strong shift of the CNP in the negative direction.  The molecular density of the second layer is considerably higher with perylene core interplanar of $< 0.34$ nm, which results in higher dipole electric field intensity than in  case of the layer comprising the flat-lying molecules.
